Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-18 thesis/chapters/ch_introduction.tex
Updates to nucleosynthesis section

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.75\linewidth]{./figures/intro/periodic_table}
  \caption{The periodic table of elements colored by the relative fraction of the nucleosynthetic processes that naturally occur in the Universe that produce each element. \textbf{Image Credit:} A modified version of the table created by Anna Frebel (\url{https://www.dropbox.com/sh/ib3agt2bkxqqk2u/AACbEw1warp74qzUh2AcWqfja?dl=0}), accessed on 03-18-19.}
\end{figure}

A variety of nucleosynthetic processes occur within stars and during their deaths that lead to the production of all of the elements of the periodic table. This is illustrated well in Figure~\ref{intro:fig:periodic table}, which shows the periodic table colored by the fractional importance of the nucleosynthetic processes that create each element. Due to the diversity in production sources across the periodic table, multi-element metal abundance measurements in stars can be used to to better understand and constrain how and where each metal is produced. Exactly which reactions occur where depend strongly on the initial abundances of elements in stars, their stellar masses, and, for heavy nuclei, the neutron density.

This means that elements from different nucleosynthetic channels are produced on different timescales during a galaxy's evolution, as set by the lifetimes of the stars in which they occur, and are released into the galaxy at different energies, with different spatial distributions within the galaxy, and at different rates. For example, O, an $\Alpha$-element is produced predominately in massive stars and is therefore released via core collapse supernovae on timescales of $\sim$~10~Myr after a star formation event, near their stellar birth sites. By contrast, Sr is produced predominately via the s-process (discussed more below) which occurs in the asymptotic giant branch (AGB) winds of lower mass stars ($M < 8$~M$_{\odot}$) at much lower energies and on timescales of $\sim 100$~Myr - $1$~Gyr. By this time, these stars would have diffused farther away from their birth sites, potentially sampling a qualitatively different portion of the ISM than earlier core collapse SNe. Generally the differences in how different timescales affect observed stellar abundances patterns are well understood and easily modelled.\footnote{The canonical example of this is the ``knee" feature in a $[\Alpha/$Fe] vs. [Fe/H] diagram, where the location of this knee signifies the point at which Type Ia enrichment (which produces significant Fe, but little to no $\Alpha$ elements) begins to have a significant contribution over core collapse SNe (rich in $\Alpha$ elements, but not Fe) \citep[e.g.[]{MatteucciBrocato1990, Geisler2005, Hill2018}. The location of this knee shifts in a galaxy depending on its SFH.} However, it is not well understood how differences in where and with what energies these events occur affect observed stellar abundance patterns. Investigating these differences is a key motivation of this dissertation.

As relevant to this work, we briefly summarize a few of the important nucleosynthetic channels shown in Figure~\ref{intro:fig:periodic table} and their sites. $\Alpha$-elements, notably O, Mg, Si, Ca, and Ti, are produced during the lifetime of massive stars (O, Mg) or during the core collapse supernova explosion itself. In either case, these elements are
